Uncertainty-aware graph-based multimodal remote sensing classification

### 2021-07-06 latex/CEUR/paper.tex
Cut down section describe graph convolutional network and simply left reference to Kipf and Welling for more details.

\section{Networks}
\label{sec::networks}
\subsection{Graph convolutional networks}
The problem involves a graph with $N$ nodes described by an adjacency matrix $\mathbf{A}\in\mathbb{R}^{N\times N}$ and node features $\mathbf{X}\in\mathbb{R}^{N\times C}$, where $C$ denotes the number of channels in the input describing each node.
The graph's degree matrix, $\mathbf{D}\in\mathbb{R}^{N\times N}$, is a diagonal matrix with elements given by $\mathbf{D}_{ii}=\sum_j\mathbf{A}_{ij} $.
For brevity, the tilde operator is used to represent the inclusion of self-connection edges in the graph, i.e. $\tilde{\mathbf{A}}=\mathbf{A}+\mathbf{I}$ and $\tilde{\mathbf{D}}=\mathbf{D}+\mathbf{I}$.
Using the first-order approximation to spectral graph convolution derived by \cite{Kipf2017}, the graph convolutional layer is given by
    \mathbf{Z}^{(l+1)}=\sigma\left(\tilde{\mathbf{D}}^{-\frac{1}{2}}\tilde{\mathbf{A}}\tilde{\mathbf{D}}^{-\frac{1}{2}}\mathbf{Z}^{(l)}\boldsymbol{W}^{(l)}\right),
\label{eq::gcn_layer}
where $\mathbf{Z}^{(l)}$, $\mathbf{Z}^{(l+1)}$ and $\boldsymbol{W}^{(l)}$ are the inputs, outputs and weights of the $\textrm{l}^\textrm{th}$ layer respectively, and $\sigma()$ is a non-linear activation function.
For the first layer, the weights are $\boldsymbol{W}^{(0)}\in\mathbb{R}^{C\times F}$ and $\mathbf{Z}^{(0)} = \mathbf{X}\in\mathbb{R}^{N\times C}$ where $F$ is the number of features in the layer's output (a hyperparameter of the layer).
Models referred to as graph convolutional networks (GCNs) from henceforth will refer to a model which consists of two of these layers used in sequence, the first using the rectified linear unit (ReLU) activation function, the second using a softmax activation.
\subsection{Subjective models}
In order to obtain models which will give node-level Dirichlet distribution parameters, the softmax output activation function used in a GCN is substituted for a ReLU function.
In this way, the model is trained to output non-negative evidence contributions, $\mathbf{E}\in\mathbb{R}^{N\times K}$, where $\mathbf{E}_{i} = \boldsymbol{\alpha}_{i}-\mathbf{1}$ and $\boldsymbol{\alpha}_{i}$ refers to the K-dimensional concentration parameters of the $\text{i}^\text{th}$ node.
In order to train an evidential model, the loss function is made up of two components; a squared error term which is minimised in order to classify a greater proportion of the nodes correctly and a variance term which is minimized to incentivize the model to provide confident predictions where possible.
This loss, $\mathcal{L}(\boldsymbol{\theta})$, is given by
\begin{split}
    \mathcal{L}(\boldsymbol{\theta})&=\sum_{i\in\mathbb{L}}\sum_k \left[(p_{ik}-y_{ik})^2+\text{Var}(p_{ik})\right],\\
    &=\sum_{i\in\mathbb{L}}\sum_k \left[(p_{ik}-y_{ik})^2+\frac{\left(\frac{\alpha_{ik}}{S_i}\right)\left(1-\frac{\alpha_{ik}}{S_i}\right)}{S_i-K}\right],
\end{split}
\label{eq:core_loss}
where $i\in\mathbb{L}$ refers to the fact that the loss is computed using a sum only over nodes in the training set, $\mathbb{L}$.
Models trained with such an output activation and loss function will be denoted using the `S-' prefix in order to indicate they provide subjective predictions, e.g. S-GCN.
\subsection{Dropout inference as a Bayesian approximation}
\label{sec::dropout_bayesian}
It has been shown that a model with dropout can be considered to represent the posterior distribution over models \cite{Gal2016}.
The dropouts are used to represent stochasticity in the model parameters \cite{Ryu2019}.
A given combination of dropouts can be considered to represent a model sampled from the approximate posterior, $q(\boldsymbol{\theta})$, with the actual posterior distribution of models being $p(\boldsymbol{\theta}|\mathcal{D})$.
Monte Carlo sampling can be used to approximately compute quantities such as the term $\mathbb{E}_{p\left(\boldsymbol{\theta}\vert \mathcal{D}\right)}\left[\mathcal{H}(p(y\vert x; \boldsymbol{\theta}))\right]$ in (\ref{eq::epistemic}) allowing the aleatoric and epistemic uncertainty to be calculated.
This can be done using the approximation 
    \mathbb{E}_{p\left(\boldsymbol{\theta}\vert \mathcal{D}\right)}\left[\mathcal{H}(p(y\vert x; \boldsymbol{\theta}))\right]\approx\frac{1}{M}\sum_{m=1}^M \mathcal{H}\left(p(y\vert x; \boldsymbol{\theta}^m)\right),
where $\boldsymbol{\theta}^m\sim q(\boldsymbol{\theta})$ is a model sampled by performing dropout inference.
When this technique has been used, the model name represents this using a `B' in its name, e.g. S-BGCN, showing that this Bayesian approximation has been used.
The Drop-GCN model, which is used for benchmarking the subjective methods, refers to the same technique when applied to a `vanilla' GCN.

>>>>

\section{Networks}
\label{sec::networks}
The problem involves a graph with $N$ nodes described by an adjacency matrix $\mathbf{A}\in\mathbb{R}^{N\times N}$ and node features $\mathbf{X}\in\mathbb{R}^{N\times C}$, where $C$ denotes the number of channels in the input describing each node.
The graph's degree matrix, $\mathbf{D}\in\mathbb{R}^{N\times N}$, is a diagonal matrix with elements given by $\mathbf{D}_{ii}=\sum_j\mathbf{A}_{ij} $.

The graph convolutional networks used are of the form proposed by \cite{Kipf2017}, where the graph convolutional layer is given by
\begin{equation}
    \mathbf{Z}^{(l+1)}=\sigma\left(\tilde{\mathbf{D}}^{-\frac{1}{2}}\tilde{\mathbf{A}}\tilde{\mathbf{D}}^{-\frac{1}{2}}\mathbf{Z}^{(l)}\boldsymbol{W}^{(l)}\right),
\label{eq::gcn_layer}
\end{equation}
where $\mathbf{Z}^{(l)}$, $\mathbf{Z}^{(l+1)}$ and $\boldsymbol{W}^{(l)}$ are the inputs, outputs and weights of the $\textrm{l}^\textrm{th}$ layer respectively, and $\sigma()$ is a non-linear activation function.
For brevity, the tilde operator is used to represent the inclusion of self-connection edges in the graph, i.e. $\tilde{\mathbf{A}}=\mathbf{A}+\mathbf{I}$ and $\tilde{\mathbf{D}}=\mathbf{D}+\mathbf{I}$.

\subsection{Subjective models}
In order to obtain models which will give node-level Dirichlet distribution parameters, the softmax output activation function used in a GCN is substituted for a ReLU function.
In this way, the model is trained to output non-negative evidence contributions, $\mathbf{E}\in\mathbb{R}^{N\times K}$, where $\mathbf{E}_{i} = \boldsymbol{\alpha}_{i}-\mathbf{1}$ and $\boldsymbol{\alpha}_{i}$ refers to the K-dimensional concentration parameters of the $\text{i}^\text{th}$ node.
In order to train an evidential model, the loss function is made up of two components; a squared error term which is minimised in order to classify a greater proportion of the nodes correctly and a variance term which is minimized to incentivize the model to provide confident predictions where possible.
This loss, $\mathcal{L}(\boldsymbol{\theta})$, is given by
\begin{equation}
\begin{split}
    \mathcal{L}(\boldsymbol{\theta})&=\sum_{i\in\mathbb{L}}\sum_k \left[(p_{ik}-y_{ik})^2+\text{Var}(p_{ik})\right],\\
    &=\sum_{i\in\mathbb{L}}\sum_k \left[(p_{ik}-y_{ik})^2+\frac{\left(\frac{\alpha_{ik}}{S_i}\right)\left(1-\frac{\alpha_{ik}}{S_i}\right)}{S_i-K}\right],
\end{split}
\label{eq:core_loss}
\end{equation}
where $i\in\mathbb{L}$ refers to the fact that the loss is computed using a sum only over nodes in the training set, $\mathbb{L}$.
Models trained with such an output activation and loss function will be denoted using the `S-' prefix in order to indicate they provide subjective predictions, e.g. S-GCN.
\subsection{Dropout inference as a Bayesian approximation}
\label{sec::dropout_bayesian}
It has been shown that a model with dropout can be considered to represent the posterior distribution over models \cite{Gal2016}.
The dropouts are used to represent stochasticity in the model parameters \cite{Ryu2019}.
A given combination of dropouts can be considered to represent a model sampled from the approximate posterior, $q(\boldsymbol{\theta})$, with the actual posterior distribution of models being $p(\boldsymbol{\theta}|\mathcal{D})$.
Monte Carlo sampling can be used to approximately compute quantities such as the term $\mathbb{E}_{p\left(\boldsymbol{\theta}\vert \mathcal{D}\right)}\left[\mathcal{H}(p(y\vert x; \boldsymbol{\theta}))\right]$ in (\ref{eq::epistemic}) allowing the aleatoric and epistemic uncertainty to be calculated.
This can be done using the approximation 
\begin{equation}
    \mathbb{E}_{p\left(\boldsymbol{\theta}\vert \mathcal{D}\right)}\left[\mathcal{H}(p(y\vert x; \boldsymbol{\theta}))\right]\approx\frac{1}{M}\sum_{m=1}^M \mathcal{H}\left(p(y\vert x; \boldsymbol{\theta}^m)\right),
\end{equation}
where $\boldsymbol{\theta}^m\sim q(\boldsymbol{\theta})$ is a model sampled by performing dropout inference.
When this technique has been used, the model name represents this using a `B' in its name, e.g. S-BGCN, showing that this Bayesian approximation has been used.
The Drop-GCN model, which is used for benchmarking the subjective methods, refers to the same technique when applied to a `vanilla' GCN.
